---
title: "Thiomon Data Analysis Report"
author: "Anonymous"
output: 
  pdf_document: 
    toc: yes
    toc_depth: 5
bibliography: references.bib 
editor_options: 
  chunk_output_type: console
---

```{r include=FALSE}
suppressPackageStartupMessages(library(aaltobda))
suppressPackageStartupMessages(library(rstan))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(loo))
suppressPackageStartupMessages(library(bayesplot))
suppressPackageStartupMessages(library(gridExtra))

```

```{r include=FALSE}
SEED = 42
library(aaltobda)
library(rstan)
library(dplyr)
library(ggplot2)
library(loo)
library(bayesplot)
library(gridExtra)
library(grid)
library(thiomon)
setwd("C:/Users/nguye/OneDrive - Aalto University/University/First year/Aalto_University_course/CS-E5710 - Bayesian Data Analysis D/bayes-project-main")
```


\newpage

# 1. Introduction


# 2. Data: Thiomon Dataset



## 2.1. Data Preprocessing 


# 3. Models
 

# 4. Priors


## 4.1. Separate model


## 4.2. Hierarchical model

# 5. Fitting the Stan models


## 5.1. Separate model  

## 5.2. Hierarchical model  


# 6. Convergence diagnostics


## 6.1. $\widehat{R}$

## 6.2. Effective sample size


## 6.3. Divergent transitions  


## 6.4. Conclusion about convergence


## 6.5. Improving convergence


# 7. Posterior predictive checks 


# 8. Model comparison with LOO-CV

# 9. Predictive performance assessment

# 10. Sensitivity analysis

## 10.1. LOO-CV
### 10.1.1. Separate model

### 10.1.2. Hierarchical model

## 10.2. Posterior distributions

# 11. Discussion of issues and potential improvements



# 13. Self-reflection


# Appendix
## A. Stan code
### A.1. Separate model

### A.2. Hierarchical model

## B. Loaded packages


# References